---
title: "Bodyfat PCR & Stepwise selection"
author: "ShuweiLiu"
date: '2022-10-07'
output: html_document
---


```{r}
# KMO Kaiser-Meyer-Olkin Measure of Sampling Adequacy
kmo = function( data ){
  
  library(MASS)
  X <- cor(as.matrix(data))
  iX <- ginv(X)
  S2 <- diag(diag((iX^-1)))
  AIS <- S2%*%iX%*%S2                      # anti-image covariance matrix
  IS <- X+AIS-2*S2                         # image covariance matrix
  Dai <- sqrt(diag(diag(AIS)))
  IR <- ginv(Dai)%*%IS%*%ginv(Dai)         # image correlation matrix
  AIR <- ginv(Dai)%*%AIS%*%ginv(Dai)       # anti-image correlation matrix
  a <- apply((AIR - diag(diag(AIR)))^2, 2, sum)
  AA <- sum(a)
  b <- apply((X - diag(nrow(X)))^2, 2, sum)
  BB <- sum(b)
  MSA <- b/(b+a)                        # indiv. measures of sampling adequacy
  
  AIR <- AIR-diag(nrow(AIR))+diag(MSA)  # Examine the anti-image of the
  # correlation matrix. That is the
  # negative of the partial correlations,
  # partialling out all other variables.
  
  kmo <- BB/(AA+BB)                     # overall KMO statistic
  
  # Reporting the conclusion
  if (kmo >= 0.00 && kmo < 0.50){
    test <- 'The KMO test yields a degree of common variance
unacceptable for FA.'
  } else if (kmo >= 0.50 && kmo < 0.60){
    test <- 'The KMO test yields a degree of common variance miserable.'
  } else if (kmo >= 0.60 && kmo < 0.70){
    test <- 'The KMO test yields a degree of common variance mediocre.'
  } else if (kmo >= 0.70 && kmo < 0.80){
    test <- 'The KMO test yields a degree of common variance middling.'
  } else if (kmo >= 0.80 && kmo < 0.90){
    test <- 'The KMO test yields a degree of common variance meritorious.'
  } else {
    test <- 'The KMO test yields a degree of common variance marvelous.'
  }
  
  ans <- list(  overall = kmo,
                report = test,
                individual = MSA,
                AIS = AIS,
                AIR = AIR )
  return(ans)
  
}    # end of kmo()
```


Reference:

https://www.statisticshowto.com/kaiser-meyer-olkin/

https://stat.ethz.ch/pipermail/r-help/2007-August/138049.html

We use KMO to test how suited our data is for Factor Analysis.



```{r}
library(car)
#setWD for wherever bodyfat csv is on machine 
body_fat <- read.csv("BodyFat.csv")
#convert to KGs
body_fat$WEIGHT <- (body_fat$WEIGHT/2.2)
#convert to CMs
body_fat$HEIGHT <- body_fat$HEIGHT * 2.54

predictors<-body_fat[4:17]
```

From our intuition, one's circumferences of different body parts will increase together when becoming fatter. So there may be strong correlation between our predictors. Thus, we can consider the principal component regression.

From the picture, we can check the correlation between our predictors.
```{r}
#pairs(predictors)
```

```{r}
kmo(predictors)
```

Check the results of KMO and take the predictors whose KMO is larger than 0.8 (in kmo(predictors)$individual).

```{r}
pred<-predictors[c(2,4:11)]
pcr_results<-prcomp(pred)
```

```{r}

#new precditors

#pcr_results$x

```

```{r}
body_fat_pcr<-cbind(body_fat$BODYFAT,predictors$AGE,predictors$HEIGHT,pcr_results$x)
colnames(body_fat_pcr)<-c("BODYFAT","AGE","HEIGHT","PC1","PC2","PC3","PC4","PC5","PC6","PC7","PC8","PC9")
body_fat_pcr<-data.frame(body_fat_pcr)
```

Stepwise selection:

```{r}
# Fit the full model 
full.model <- lm(BODYFAT ~., data = body_fat_pcr)
# Stepwise regression model
step.model <- stepAIC(full.model, direction = "both", 
                      trace = FALSE)
summary(step.model)
```




