---
title: "Body Fat Analysis"
author: "Marwan Lloyd, Shuwei Liu, David Gao"
date: '2022-10-07'
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
library(car)
library(ggplot2)
#setWD for wherever bodyfat csv is on machine 
#setwd("~/Desktop/UW-Madison/Stat 628/Module 2 Starter Materials")
body_fat <- read.csv('adjusted_bodyfat.csv')
```
Stepwise Part

```{r}
library(MASS)
# Fit the full model 
full.model2 <- lm(BODYFAT ~AGE+WEIGHT+HEIGHT+ADIPOSITY+NECK+CHEST+ABDOMEN+HIP+THIGH+KNEE+ANKLE+BICEPS+FOREARM+WRIST, data = body_fat)
# Stepwise regression model
stepwise.model <- stepAIC(full.model2, direction = "both", 
                      trace = FALSE)
summary(stepwise.model)
```

PCA+Stepwise Part

```{r}
# KMO Kaiser-Meyer-Olkin Measure of Sampling Adequacy
kmo = function( data ){
  
  library(MASS)
  X <- cor(as.matrix(data))
  iX <- ginv(X)
  S2 <- diag(diag((iX^-1)))
  AIS <- S2%*%iX%*%S2                      # anti-image covariance matrix
  IS <- X+AIS-2*S2                         # image covariance matrix
  Dai <- sqrt(diag(diag(AIS)))
  IR <- ginv(Dai)%*%IS%*%ginv(Dai)         # image correlation matrix
  AIR <- ginv(Dai)%*%AIS%*%ginv(Dai)       # anti-image correlation matrix
  a <- apply((AIR - diag(diag(AIR)))^2, 2, sum)
  AA <- sum(a)
  b <- apply((X - diag(nrow(X)))^2, 2, sum)
  BB <- sum(b)
  MSA <- b/(b+a)                        # indiv. measures of sampling adequacy
  
  AIR <- AIR-diag(nrow(AIR))+diag(MSA)  # Examine the anti-image of the
  # correlation matrix. That is the
  # negative of the partial correlations,
  # partialling out all other variables.
  
  kmo <- BB/(AA+BB)                     # overall KMO statistic
  
  # Reporting the conclusion
  if (kmo >= 0.00 && kmo < 0.50){
    test <- 'The KMO test yields a degree of common variance
unacceptable for FA.'
  } else if (kmo >= 0.50 && kmo < 0.60){
    test <- 'The KMO test yields a degree of common variance miserable.'
  } else if (kmo >= 0.60 && kmo < 0.70){
    test <- 'The KMO test yields a degree of common variance mediocre.'
  } else if (kmo >= 0.70 && kmo < 0.80){
    test <- 'The KMO test yields a degree of common variance middling.'
  } else if (kmo >= 0.80 && kmo < 0.90){
    test <- 'The KMO test yields a degree of common variance meritorious.'
  } else {
    test <- 'The KMO test yields a degree of common variance marvelous.'
  }
  
  ans <- list(  overall = kmo,
                report = test,
                individual = MSA,
                AIS = AIS,
                AIR = AIR )
  return(ans)
  
}    # end of kmo()
```


Reference:

https://www.statisticshowto.com/kaiser-meyer-olkin/

https://stat.ethz.ch/pipermail/r-help/2007-August/138049.html

We use KMO to test how suited our data is for Factor Analysis.



```{r}
predictors<-body_fat[3:16]
```

From our intuition, one's circumferences of different body parts will increase together when becoming fatter. So there may be strong correlation between our predictors. Thus, we can consider the principal component regression.

From the picture, we can check the correlation between our predictors.

```{r}
kmo(predictors)
```

Check the results of KMO and take the predictors whose KMO is larger than 0.8 (in kmo(predictors)$individual).

```{r}
pred<-predictors[c(2,4:11)]
pcr_results<-prcomp(pred)
```


```{r}
body_fat_pcr<-cbind(body_fat$BODYFAT,predictors$AGE,predictors$HEIGHT,pcr_results$x)
colnames(body_fat_pcr)<-c("BODYFAT","AGE","HEIGHT","PC1","PC2","PC3","PC4","PC5","PC6","PC7","PC8","PC9")
body_fat_pcr<-data.frame(body_fat_pcr)
```

Stepwise selection:

```{r}
# Fit the full model 
full.model <- lm(BODYFAT ~., data = body_fat_pcr)
# Stepwise regression model
PCAstep.model <- stepAIC(full.model, direction = "both", 
                      trace = FALSE)
summary(PCAstep.model)
```

```{r}
sort(unique(body_fat$BODYFAT)) # <- from this we see some body fat numbers make no sense
```

```{r}
#remove ID column remove density column, and remove adiposity as that's weight/height and keeping it could lead to multicollinearity issues
body_fat_lm <- lm(body_fat$BODYFAT ~ . , body_fat)
summary(body_fat_lm)
#Check the residual plot to see if it is a good fit
res.full = resid(body_fat_lm)
plot(body_fat$BODYFAT,res.full, ylab = "Residuals", xlab = "Bodyfat",main = "Residual plot for full model")
#turns out not a good fit, overfitting
```

```{r}
#Get VIF's
# VIFs start at 1 and have no upper limit. A value of 1 indicates that there is no correlation between this independent variable and any others. VIFs between 1 and 5 suggest that there is a moderate correlation, but it is not severe enough to warrant corrective measures. VIFs greater than 5 represent critical levels of multicollinearity where the coefficients are poorly estimated, and the p-values are questionable.
vif(body_fat_lm)
```

```{r}
#remove high p-value variables (higher than .2)
#so  remove height, chest, knee, ankle, biceps
#remaining variables are: age, weight, neck, abdomen, hip, thigh, forearm, wrist
bf_2 <- body_fat[c(1,3,4,7,9,10,11,15,16)]
fit_2 <- lm(BODYFAT ~ . , bf_2)
summary(fit_2)
#check the residual plot, turns out a better fit
res.full.drop = resid(fit_2)
plot(bf_2$BODYFAT,res.full.drop, ylab = "Residuals", xlab = "Bodyfat",main = "Residual plot for full model after dropping")
```

```{r}
#proposing models
#for simplicity this model is good as it's only 3 variables that are easy to get 
#Is better than just Age + Weight as it's R^2 is .3 higher 
#R^2 is only .02 less than fit_2's R^2
proposinglm <- lm(BODYFAT ~ AGE + WEIGHT + ABDOMEN, bf_2)
summary(proposinglm)
res.proposing = resid(proposinglm)
plot(bf_2$BODYFAT,res.proposing, ylab = "Residuals", xlab = "Bodyfat",main = "Residual plot for proposing model after dropping")
```

```{r}
#proposing model 2 
#simpler but loses a lot of accuracy by missing one variable
proposinglm1 <- lm(BODYFAT ~ AGE + WEIGHT, bf_2)
summary(proposinglm1)
res.proposing1 = resid(proposinglm1)
plot(bf_2$BODYFAT,res.proposing1, ylab = "Residuals", xlab = "Bodyfat",main = "Residual plot for second proposing model after dropping")
```

```{r}
#Model w/ 3 most statistically significant values but is less accurate than Age + weight + abdomen
proposinglm2 <- lm(BODYFAT ~ FOREARM + WRIST + ABDOMEN, body_fat)
summary(proposinglm2)
res.proposing2 = resid(proposinglm2)
plot(bf_2$BODYFAT,res.proposing2, ylab = "Residuals", xlab = "Bodyfat",main = "Residual plot for 3 predictors model after dropping")
```

```{r}
#proposing model 4
#Seems like best
#highest R^2 of proposed ones w/ .7187 
#lower standard error than AGE + WEIGHT + ABDOMEN
proposinglm3 <- lm(BODYFAT ~ WEIGHT + ABDOMEN, body_fat)
summary(proposinglm3)
res.proposing3 = resid(proposinglm3)
plot(bf_2$BODYFAT,res.proposing3, ylab = "Residuals", xlab = "Bodyfat",main = "Residual plot for 2 predictors model after dropping")
```


```{r FIRST GO}
library(shiny)
ui <- fluidPage(
selectInput("stat_dropdown", "Select a Stat of interest", choices = colnames(body_fat)[3:17], multiple = FALSE),
plotOutput("general_scatter")
)
server <- function(input, output) { 
  output$general_scatter <-  renderPlot({ 
  ggplot(body_fat,aes(input$stat_dropdown,BODYFAT)) + 
      labs(x = input$stat_dropdown, y = "Body Fat", title = "Body Fat Scatterplot") +
     geom_jitter(aes(x = input$stat_dropdown,y = BODYFAT, color = input$stat_dropdown, size = BODYFAT))
})
  }
app <- shinyApp(ui, server)
app
```


```{r REAL TRY}
library(shiny)
library(shinythemes)
ui <- 
  navbarPage("Group 2 Shiny", collapsible = TRUE, inverse = TRUE, theme = shinytheme("spacelab"),
#             tabPanel("All Stat Scatterplots",
#                      fluidPage(
#                      selectInput("stat_dropdown", "Select a Stat of Interest", #choices = colnames(body_fat)[3:17], multiple = FALSE),
#                      plotOutput("general_scatter")
#                      )),
             tabPanel("Model Calculator",
                      tags$p("The model that our project decided to use was a multi-linear regression of bodyfat percentage on two variables: weight and abdomen circumfrence. The final equation (used in this calculator) determined for this regression was:"), tags$h3("Body Fat % = -40.47 - .31*Weight + .91 * Abdomen"), tags$h4("---------------------------------------------------------"),
                      sidebarLayout(sidebarPanel(numericInput("Weight", "Weight (in kg):", 0, min = 0, step = .01), numericInput("Abdomen", "Abdomen Circumfrence (in cm):", 0, min = 0, step = .01)),
                        mainPanel("Calculated Bodyfat:", textOutput("calculated_bodyfat"), tags$head(tags$style("#calculated_bodyfat{color: black;
                                 font-size: 20px;
                                 font-style: bold;
                                 }"
                         )
              )))
                      ),
             tabPanel("About", tags$p("This shiny app is maintained by Marwan Lloyd '23. For any questions, concerns, or feedback please reach out him at either of the below emails:"), tags$address("Marwan.Lloyd@Wisc.edu"), tags$address("Melloyd2@Wisc.edu"))
)
server <- function(input, output) { 
  
  observeEvent(input$Weight,{
    if(input$Weight < 0 || isTRUE(is.na(input$Weight)) ){ 
      x <- 0}
    else{
      x <- input$Weight
    }
    updateNumericInput(getDefaultReactiveDomain(), "Weight", value = round(x,digits = 2))})
  
  observeEvent(input$Abdomen,{
    if(input$Abdomen < 0 || isTRUE(is.na(input$Abdomen)) ){ 
      y <- 0}
    else{
      y <- input$Abdomen
    }
    updateNumericInput(getDefaultReactiveDomain(), "Abdomen", value = round(y,digits = 2))})
  
  output$calculated_bodyfat <- renderText(paste(c(round(-40.47 - .31*input$Weight + .91 * input$Abdomen,2)),"%"))
  
  output$general_scatter <-  renderPlot({ 
  ggplot(body_fat,aes(input$stat_dropdown,BODYFAT)) + 
      labs(x = input$stat_dropdown, y = "Body Fat", title = "Body Fat Scatterplot") +
     geom_jitter(aes(x = input$stat_dropdown,y = BODYFAT, color = input$stat_dropdown, size = BODYFAT))
})
  }
shinyApp(ui = ui, server = server)
```
